\documentclass[12pt,a4paper]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{tocloft}
\usepackage{lipsum}
\usepackage{enumitem}
\usepackage{array}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}

% Page setup
\pagestyle{fancy}
\fancyhf{}
\rhead{Assignment 2 - Trusted Computing}
\lhead{TPM and TEE}
\rfoot{Page \thepage}

% Listings setup for code
\lstset{
    basicstyle=\ttfamily\small,
    backgroundcolor=\color{gray!10},
    frame=single,
    breaklines=true,
    columns=fullflexible,
    keepspaces=true,
    showstringspaces=false,
    numbers=left,
    numberstyle=\tiny\color{gray},
    commentstyle=\color{green!60!black},
    keywordstyle=\color{blue},
    stringstyle=\color{orange},
}

% Hyperref setup
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
    pdftitle={Assignment 2 - Trusted Computing},
}

% Title formatting
\titleformat{\section}
  {\normalfont\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}
  {\normalfont\large\bfseries}{\thesubsection}{1em}{}

% Document begins
\begin{document}

% ============================================================
% TITLE PAGE
% ============================================================

\begin{titlepage}
    \centering
    \vspace{2cm}

    {\huge \textbf{Assignment 2}}\\[0.5cm]
    {\LARGE Trusted Computing}\\[0.3cm]
    {\Large TPM Architecture and TEE}\\[2cm]

    \rule{\linewidth}{0.5mm}\\[0.4cm]
    {\Large \textbf{Theory and Practical Implementation}}\\[0.4cm]
    \rule{\linewidth}{0.5mm}\\[2cm]

    \begin{tabular}{rl}
        \textbf{Student Name:} & Yuvan Raj Krishna \\[0.3cm]
        \textbf{Register Number:} & 22011102127 \\[0.3cm]
        \textbf{Course:} & Trusted Computing \\[0.3cm]
        \textbf{Date:} & November 3, 2025 \\
    \end{tabular}

\end{titlepage}

% ============================================================
% TABLE OF CONTENTS
% ============================================================

\newpage
\tableofcontents
\newpage

% ============================================================
% PART A - THEORY QUESTIONS
% ============================================================

\part{Theory Questions (10 marks)}

% Include all Part A content (CONCISE VERSIONS)
\input{parta_q1_concise.tex}
\input{parta_q2_concise.tex}
\input{parta_q3_concise.tex}
\input{parta_q4_concise.tex}

% ============================================================
% PART B - PRACTICAL TASKS
% ============================================================

\newpage
\part{Practical Tasks (10 marks)}

\input{partb_concise.tex}

% ============================================================
% REFERENCES
% ============================================================

\newpage
\section*{References}
\addcontentsline{toc}{section}{References}

\begin{enumerate}[leftmargin=*]
    \item Trusted Computing Group, ``TPM 2.0 Library Specification'', Part 1: Architecture, Rev. 1.59, November 2019

    \item Arthur, W., \& Challener, D. (2015). \textit{A Practical Guide to TPM 2.0}. Apress

    \item TCG PC Client Platform Firmware Profile Specification, Family 2.0, Level 00 Revision 1.05

    \item Sailer, R., et al. (2004). ``Design and Implementation of a TCG-based Integrity Measurement Architecture''. \textit{USENIX Security Symposium}

    \item ARM, ``ARM Security Technology - Building a Secure System using TrustZone Technology'', 2022

    \item Ngabonziza, B., et al. (2016). ``TrustZone Explained: Architectural Features and Use Cases''. \textit{IEEE Cybersecurity Development Conference}

    \item Costan, V., \& Devadas, S. (2016). ``Intel SGX Explained''. \textit{IACR Cryptology ePrint Archive}

    \item McKeen, F., et al. (2013). ``Innovative Instructions and Software Model for Isolated Execution''. \textit{HASP Workshop}

    \item GlobalPlatform, ``TEE Internal Core API Specification'', Version 1.3.1, 2022

    \item Intel Corporation, ``Intel SGX Developer Reference'', 2024
\end{enumerate}

\end{document}
